\papertitle

### 2014-03-18 final-report/paper.tex
added figures, discussion etc.

The \graph view allows the user to selectively browse different performance visualizations tied to a given query execution. The \graph view provides as the name indicates, a visualization of the physical query plan. A query plan is an ordered set of steps used to access data in database management systems. The physical query plan is represented by a graph where each node represents a query fragment, and each link represents inter-worker communication from one query fragment to the next. Each query fragment is a collection of query operators. One can think of a query fragment as a job unit that gets scheduled on a cluster of machines. Consequently transitioning from one query fragment to another generally leads to all-to-all communication between worker nodes on the cluster.

@@

\textbf{Graph Rendering:} We use D3 \cite{d3} to render the \graph graph, which allows us to support various interactions and transistions. We use GraphViz \cite{Ellson01graphviz} in the backend to generate graph layout information. GraphViz allows us to generate graph layouts that are optimized to minimize area footprint. The layout information is then fed into a D3-based rendering engine which supports various interactions and animations.

\textbf{Graph Navigation:} A query plan can be arbitrarily large. Thus we offer two mechanisms that facilitate exploration of the graph for the user: (1) expanding/reducing fragments and (2) zooming. The first mechanism can reduce the size of a graph by a constant factor by collapsing operators that compose a fragment into a single fragment node. The user can click once on a collapsed fragment to expand it, and click once on an expanded fragment to collapse it. Expanding or reducing a node can cause large changes in the graph layout as GraphViz changes the graph layout to minimize overall area. To address these layout changes, we implemented transitions to allow the user to track the fragments as those get reshuffled. The second mechanism that facilitates exploration of the graph for the user is zooming. This feature was implemented in D3 and allows the user to quickly jump from one part of the graph to the next, by zooming in and out of the graph view.

@@

The resulting \graph visualization is displayed in Figure~\ref{fig:TODO}. The user will navigate the \graph visualization in the browser panel in the next case studies to identify several performance symptoms.

@@

\subsection{Case 3: Identifying Execution Skew using the \fragment View}
The \fragment visualization allows the user to dive down into query execution schedules at the worker level.

>>>>

\begin{figure}[ht]
  \centering
  \begin{subfigure}[b]{0.49\columnwidth}
    \includegraphics[width=0.8\columnwidth]{images/graph_collapsed}
    \caption{Large query plan collapsed.}
    \label{fig:graph}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\columnwidth}
    \includegraphics[width=0.8\columnwidth]{images/graph_expanded}
    \caption{Expanding Fragment 2 reveals its operators.}
    \label{fig:graph_expanded}
  \end{subfigure}
  \caption{The \graph view is used by \system to help the user browse performance visualizations.}
  \label{fig:graph}
\end{figure}

The \graph view as pictured in Figure~\ref{fig:graph} allows the user to selectively browse different performance visualizations tied to a given query execution. The \graph view provides as the name indicates, a visualization of the physical query plan. A query plan is an ordered set of steps used to access data in database management systems. The physical query plan is represented by a graph where each node represents a query fragment, and each link represents inter-worker communication from one query fragment to the next. Each query fragment is a collection of query operators. One can think of a query fragment as a job unit that gets scheduled on a cluster of machines. Consequently transitioning from one query fragment to another generally leads to all-to-all communication between worker nodes on the cluster.

@@

\textbf{Graph Rendering:} We use D3 \cite{d3} to render the \graph graph, which allows us to support various interactions and transistions. We use GraphViz \cite{Ellson01graphviz} in the backend to generate graph layout information. GraphViz allows us to generate graph layouts that are optimized to minimize area footprint. The layout information is then fed into a D3-based rendering engine which supports various interactions and animations techniques.

\textbf{Graph Navigation:} A query plan can be arbitrarily large. Thus we offer two mechanisms that facilitate exploration of the graph for the user: (1) expanding/reducing fragments and (2) zooming. The first mechanism can reduce the size of a graph by a constant factor by collapsing operators that compose a fragment into a single fragment node. The user can click once on a collapsed fragment to expand, and click once on an expanded fragment to collapse it. Expanding or reducing a node can cause large changes in the graph layout as GraphViz changes the graph layout to minimize overall area. This graph reshuffeling is illustrated in Figure~\ref{graph} where expanding Frag2 causes the layout to change. To address these layout changes, we implemented transitions to allow the user to track the fragments as those get reshuffled. The second mechanism that facilitates exploration of the graph for the user is zooming. This feature was implemented in D3 and allows the user to quickly jump from one part of the graph to the next, by zooming in and out of the graph view.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.4\columnwidth]{images/graph_join}
  \caption{This view shows the \graph for a join.}
  \label{fig:join}
\end{figure}

The resulting \graph visualization is displayed in Figure~\ref{fig:join}. The user will navigate the \graph visualization in the browser panel in the next case studies to identify several performance symptoms.

@@

\subsection{Case 3: Identifying Execution Skew using the \fragment View}

\begin{figure}[ht]
  \includegraphics[width=\columnwidth]{images/execution_skew}
  \caption{This \fragment view shows heavy execution skew on worker~5.}
  \label{fig:exec_skew}
\end{figure}

The \fragment visualization allows the user to dive down into query execution details at the worker level. The \fragment view in Figure~\ref{fig:exec_skew} was obtained by selecting Frag2, which is the fragment that performs the join, in the \graph view. The user can observe at a glance that certain workers produce much more data as a result of the join execution. Particularly, we saw in Case 2 Figure~\ref{fig:skew}, that tuples sent to Fragment 2 from Fragment 1 were heavily skewed towards being sent to worker~1 and worker~5. In Figure~\ref{fig:exec_skew}, the user can now see that as a result of the join, worker~5 ends up producing most tuples that have to be written back to disk. The color scheme used for each worker schedules in the \fragment view is color-coded to match the operators in Figure~\ref{fig:join}. Consequently, one can quickly observe in the \fragment view that worker~5 is busy sending join results from t=2.5s to t=5.0s, which differes from the rest of the workers. Using similar reasoning, one can conclude that worker 5 presents a performance bottleneck in the execution of the query on Myria. Perhaps the user can modify the way the input data set gets partitioned. Alternatively she could change the number of workers allocated for the job.

\subsection{Discussion}

All in all, \system allows the user to get much more insight into the execution of a query than what was available before, i.e. text-based performance logs, and runtimes. Being able to navigate performance data using a simple interface can indeed facilitate the task of a developper who is designing Myria's back end, or a Myria user who is writing complex queries to process large amounts of data. That said, \system can only help the user visualize query execution performance, but won't necessarily point the user to what she needs to do to improve performance of her query execution.
